The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-07-02 paper/iphas-dr2-rev1.tex
Added section 7.3

\begin{table}
    \caption{Key properties of IPHAS DR2.}
    \label{tbl:survey}
    \begin{center}
        \begin{tabular}{ll}
        \toprule
        Property & Value \\
        \midrule
        Telescope & 2.5m Isaac Newton Telescope, La Palma \\
        Instrument & Wide Field Camera \\
        Detectors & Four 2048 $\times$ 4100 pixel CCD's \\
        %Detector field of view & 11 $\times$ 23 arcmin \\
        Detector scale & 0.33 arcsec/pixel \\        
        Filters & $r$, $i$, \ha \\
        Filter properties & See Table~2 \\
        Exposure times & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
		Saturation limit & TBC \\
        Detection limit (5$\sigma$, Vega) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
        PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
        Survey area & $\sim 1860$~deg$^2$ \\
        Footprint boundaries & $ -5\degr < b < +5\degr,\ 29\degr < \ell < 215\degr$ \\
        DR2 observing period & August 2003 - November 2012 \\
        Website & www.iphas.org \\
        \bottomrule
        \end{tabular}
	\end{center}
\end{table}

@@

>>>>

\begin{table}
    \caption{Key properties of IPHAS DR2.}
    \label{tbl:survey}
    \begin{center}
        \begin{tabular}{ll}
        \toprule
        Property & Value \\
        \midrule
        Telescope & 2.5m Isaac Newton Telescope \\
        Instrument & Wide Field Camera \\
        Detectors & Four 2048 $\times$ 4100 pixel CCD's \\
        %Detector field of view & 11 $\times$ 23 arcmin \\
        Detector scale & 0.33 arcsec/pixel \\        
        Filters & $r$, $i$, \ha \\
        Filter properties & See Table~2 \\
        Exposure times & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
		Saturation limit & $\sim$13 ($r$), $\sim$12.5 ($i$), $\sim$12.5 (\ha) \\
        Detection limit (5$\sigma$) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
		Magnitude system & Vega \\
        PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
        Survey area & $\sim 1860$~deg$^2$ \\
        Footprint boundaries & $ -5\degr < b < +5\degr,\ 29\degr < \ell < 215\degr$ \\
        Observing period & August 2003 - November 2012 \\
        Website & www.iphas.org \\
        \bottomrule
        \end{tabular}
	\end{center}
\end{table}

@@

\subsection{$r - \ha$ excess as a quantitative measure of H$\alpha$ emission}

{ \bf

As well as discover emission line stars, IPHAS data can provide a
first estimate of the equivalent width of line emission.
How this is done has been discussed in previous works 
-- most notably by \citet{Drew2005}
and more recently by \cite{Barentsen2011a}.
Given that the narrowband filter used has a FWHM of 95~\AA ,
it is to be expected that the appearance
in a spectrum of line emission corresponding to an equivalent width (EW)
of 95~\AA\ would increase $r - \ha$ by 0.75 magnitudes
(i.e. the flux captured within the narrow band doubles).
Similarly, 10~\AA\ of line emission
should increase $r - H\alpha$ by 0.11~mag.

This simple reasoning breaks down
when the $H\alpha$ emission becomes so bright
that it also dominates the flux captured across the entire $r$ broadband.
It was noted by \cite{Drew2005}
that the very bright line limiting value 
of $r - \ha$ is $\sim$3.1 in the Vega system.
This fact, on its own, implies that any apparent detection
of an \ha\ emission line star without a corresponding $r$ 
detection is only real if the \ha\ magnitude is not brighter 
by more than 3.1~mag relative to the $r$ detection limit
(i.e. typically, credible H$\alpha$-only sources
must be fainter than $\sim$18th mag).

In order to infer the emission EW from the DR2 $r - \ha$ colour
it is necessary to know the value this colour would have
in the case of no excess line emission.
In the {\em general} case this is not possible
without some prior knowledge of the star's reddening and spectral type
(i.e. the inferences drawn for an unreddened M dwarf or
a reddened F star would not be the same -- the absolute value of the 
equivalent width in the latter case would be greater because the 
reddened F star {\em without} emission would lie at lower $r -\ha$).
This is the situation faced in the study of IC 1396 by
\cite{Barentsen2011a} in which \ha\ EWs were inferred
relative to a reddened sequence of K-M stars
appropriate to this star-forming region.
A different choice of reddening would have resulted
in different EW estimates.

However it was found by \cite{Drew2005}
that for earlier type stars in which the spectrum
is not heavily modulated by line features or molecular bands
the estimation of net emission EW is more straightforward and unambiguous.  
It was demonstrated, for spectral slopes ranging
from the Rayleigh-Jeans limit to that appropriate to an early G star
and for a range of reddenings,
that it is possible to draw a unique set of constant equivalent-width 
lines in the $(r - \ha, r - i)$ diagram
that can be compared directly with the positions
of candidate emission line stars,
thereby predicting emission EW \citep[see fig.~6, ][]{Drew2005}.
The EW in this case should be understood
to be the measure of the emission observed
above the interpolated continuum
(i.e. no correction is present
for any assumed infilled photospheric line absorption).

In Fig.~\ref{fig:raddi} we show how well this works in reality:
the predicted lines of constant emission EW
in the $(r - \ha, r - i)$ plane
are compared with the positions of a sample
of over 200 classical Be stars in Perseus (from Raddi et al, 2014).
Stars of this type come with the advantage that secular variation,
whilst itcertainly occurs,
is not as common as in T Tau stars.
The objects plotted are a bright selection ($r \lesssim 16.5$)
for which there are spectroscopic determinations
of \ha\ emission equivalent width.
A colour scale has been applied to the data points
that darkens as EW increases.
The gradation apparent in the shading of the datapoints
follows the trend set by the constant EW lines quite well.  
The experience has been that the EW deduced from
IPHAS photometry for individual objects is commonly within 10~\AA\
of subsequent spectroscopic measurement
\citep[see also fig.~5 of][]{Barentsen2011a}.
Evidently, the photometry is well-suited to statistical measurement
across large samples, while for individual objects a useful
approximation is delivered.

\begin{figure}
\begin{center}
  \includegraphics[width=\linewidth]{figures/raddi.pdf}
    \caption{Photometric predictions of \ha\ equivalent width (EW)
    compared with spectroscopic measurements
    of a sample of classical Be stars (from Raddi et al 2014, submitted).
    The broken lines are predicted lines
    of constant net emission EW,
    while the analogous spectroscopic EW is represented
    by the colouring of the data points.
    The unreddened main sequence, which normally serves 
    as the upper bound to the main stellar locus,
    is drawn as a solid black line.
    Strongly reddened stars with not-so-strong line emission
    will fall below it
    and be hard to pick out without additional information.
    The two vertical lines drawn are 'curves of growth'
    showing how $r - H\alpha$ increases 
    -- and $r - i$ drops a little --
    as more and more line emission is added,
    raising the flux in \ha\ narrowband
    and to a lesser proportion the flux in the $r$ band.
    The blue line represents the trend
    for an unreddened Rayleigh-Jeans continuum
    (with superposed \ha\ emission),
    while the green line is the trend
    for the case of an unreddened A0 continuum.}
    \label{fig:raddi}
\end{center}
\end{figure}
